SunPy - Python for Solar Physics

### 2014-01-27 2-1-Map.tex
all of my changes

\subsection{Map}
The purpose of the Map datatype is to represent 2D spatial data such as images of the Sun and 
inner heliosphere. It provides a wrapper around a data array and its associated
spatial coordinates and as well as meta data. The \texttt{Map} object provides 
convenience methods for many functions 
such as rotation and re-sampling as well as convenience visualisation 
functions, providing an easy to use as well as powerful interface. The \texttt{Map} object
also provides a convenient interface for loadings data from a variety of sources, typically
a FITS file as shown in Figure~\ref{code:aia_1}.

The architecture of the \texttt{Map} object consists of a template map called
\texttt{GenericMap} which is a subclass of \texttt{astropy.nddata.NDData}. \texttt{NDData} is a 
generic wrapper around a \texttt{ndarray} with a \texttt{.meta} attribute to store meta data.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does not make full
use of its capabilities but this inheritance structure provides 
for future integration of \texttt{astropy}. In order to provide instrument/
detector specific integration \textt{GenericMap} is subclassed on creation. 
Each subclass of \texttt{GenericMap} can register with the \texttt{Map} creation function (i.e. factory)
by implementing a method that returns \texttt{True} if matching meta data 
for that instrument or detector is found in the file header. The \texttt{Map} factory 
will then automatically return an instance of the specific \texttt{GenericMap} 
subclass. As of this version Sunpy has \texttt{Map} specialisations for the 
following instruments: 
\textit{YOHKOH/SXT},\textit{SOHO/EIT} and \textit{LASCO}, \textit{RHESSI}, 
\textit{STEREO/EUVI} and \textit{COR}, \textit{HINODE/XRT},
\textit{PROBA2/SWAP}, \textit{SDO/AIA} and \textit{HMI}, 
and \textit{IRIS} SJI frames.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\caption{Demonstration of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The Map is created from an \textit{AIA} FITS file and the 
key meta data and array overview is printed. Then a quick view plot is created 
by using the \texttt{.peek()} method.}
\label{code:aia_1}
\end{listing}

As well as providing the base classes the map sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Figure~\ref{code:compmap_1} and \ref{code:mapcube_1} shows how to interact 
with these classes.
In addition, Figure~\ref{code:mapcube_2} demonstrates how the \texttt{MapCube} 
classes interacts with the \texttt{animation} module to save a video.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
import matplotlib.pyplot as plt
import sunpy.map

#Create a composite map from two files
compmap = sunpy.map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
composite=True)

#Set the RHESSI image (index 1) to have contour levels and a red colour map.
compmap.set_levels(1,range(0,50,5),percent=True)
compmap.set_colors(1,'Reds_r')

#Plot the result and crop
ax = plt.subplot()
compmap.plot()
ax.axis([200,600,-600,-200])
\end{minted}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\caption{Example demonstrating a CompositeMap plot, using contours and how 
SunPy integrates with matplotlib's pyplot functional interface.}
\label{code:compmap_1}
\end{listing}

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
import sunpy.map

compmap = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
compmap.peek()
\end{minted}
\includegraphics[width=0.8\columnwidth]{aia_cube_controls}
\caption{An example showing creation of a MapCube from a glob file search. The 
resultant plot makes use of matplotlib's interactive widgets to allow scrolling 
through the MapCube.}
\label{code:mapcube_1}
\end{listing}

>>>>

\subsection{Map}
The purpose of the Map datatype is to represent 2D spatial data such as images of the Sun and 
inner heliosphere. It provides a wrapper around a data array and its associated
spatial coordinates and as well as meta data. The \texttt{Map} object provides 
convenience methods for many functions 
such as rotation and re-sampling as well as convenience visualisation 
functions, providing an easy to use as well as powerful interface. The \texttt{Map} object
also provides a convenient interface for loadings data from a variety of sources, typically
a FITS file as shown in Listing~\ref{code:aia_1}.

The architecture of the \texttt{Map} object consists of a template map called
\texttt{GenericMap} which is a subclass of \texttt{astropy.nddata.NDData}. \texttt{NDData} is a 
generic wrapper around a \texttt{ndarray} with a \texttt{.meta} attribute to store meta data.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does not make full
use of its capabilities but this inheritance structure provides 
for future integration of \texttt{astropy}. In order to provide instrument/
detector specific integration \texttt{GenericMap} is subclassed on creation. 
Each subclass of \texttt{GenericMap} can register with the \texttt{Map} creation function (i.e. factory)
by implementing a method that returns \texttt{True} if matching meta data 
for that instrument or detector is found in the file header. The \texttt{Map} factory 
will then automatically return an instance of the specific \texttt{GenericMap} 
subclass. As of this version SunPy has \texttt{Map} specialisations for the 
following instruments: 
\textit{YOHKOH/SXT},\textit{SOHO/EIT} and \textit{LASCO}, \textit{RHESSI}, 
\textit{STEREO/EUVI} and \textit{COR}, \textit{HINODE/XRT},
\textit{PROBA2/SWAP}, \textit{SDO/AIA} and \textit{HMI}, 
and \textit{IRIS} SJI frames.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Demonstration of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The Map is created from an \textit{AIA} FITS file and the 
key meta data and array overview is printed. Then a quick view plot is created 
by using the \texttt{.peek()} method.}
\label{code:aia_1}
\end{listing}

As well as providing the base classes the map sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Listing~\ref{code:compmap_1} and \ref{code:mapcube_1} shows how to interact 
with these classes.
In addition, Listing~\ref{code:mapcube_2} demonstrates how the \texttt{MapCube} 
classes interacts with the \texttt{animation} module to save a video.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import matplotlib.pyplot as plt
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
...                         composite=True)
>>> compmap.set_levels(1, range(0,50,5), percent=True)
>>> compmap.set_colors(1, 'Reds_r')
#Plot the result and crop
>>> ax = plt.subplot()
>>> compmap.plot()
>>> ax.axis([200, 600, -600, -200])
>>> plt.show()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\end{center}
\caption{Example demonstrating a CompositeMap plot, using contours and how 
SunPy integrates with matplotlib's pyplot functional interface.}
\label{code:compmap_1}
\end{listing}

%schriste - stuart, i agree if you want to remove the following two listings.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
import sunpy.map

compmap = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
compmap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_cube_controls}
\end{center}
\caption{An example showing creation of a MapCube from a glob file search. The 
resultant plot makes use of matplotlib's interactive widgets to allow scrolling 
through the MapCube.}
\label{code:mapcube_1}
\end{listing}

### 2014-01-27 2-Data_Types.tex
added bg to listings in visualisation section
all of my changes

\section{Core Data Types}
One of the primary focus' of the SunPy library is to provide data structures 
that are specifically designed for the various types of data generally 
processed for solar physics. Currently SunPy provides three data 
types, each separate from one another to process 2D spatial data 
(\texttt{Map}), 1D temporal series (\texttt{LightCurve}) and 1 and 2D spectra 
(\texttt{Spectrum} and \texttt{Spectra}).

The design of SunPy's core data types is as a wrapper around a data object, 
either a \texttt{numpy} \texttt{ndarray} or a \texttt{pandas} 
\texttt{DataFrame}, for \texttt{Map} and \texttt{Spectra} and 
\texttt{LightCurve} respectively. Each data type has a \texttt{.data} 
attribute where this data object is located and some meta data. It is 
possible to instantiate the data types from various different 
sources, \textit{e.g.} files, urls, arrays or time ranges. The 
specific sources supported is dependant upon which data type and 
which instrument. All the SunPy data types support specialisation on 
a per-instrument basis, \textit{i.e.} \texttt{Map} has a 
\texttt{AIAMap} sub-type. This design allows simple extensions to 
cope with the diversity of data from different sources in solar 
physics. As well as support for different instruments and data 
sources each data type supports different visualisation wrappers that 
allow both quick view plotting (the peek command) and more 
fine-grained control plotting that integrates well with 
\texttt{matplotlib} (the plot command).

@@

\subsection{Visualisation}
\label{subsec:Viz}
As has been demonstrated in this section, the core SunPy datatypes 
include visualisation code that is specialised to that data type. 
These visualisation methods all currently utilise the matplotlib 
package, and are designed in such a way that they integrate well with 
the pyplot functional interface of matplotlib, as demonstrated in 
Listing \ref{code:mpl_1}.

\begin{listing}[H]
\begin{minipage}{0.49\columnwidth}
\begin{minted}{python}
import sunpy.map
import matplotlib.pyplot as plt

#Read in a aia file using SunPy
aiamap = sunpy.map.Map('aia.fits')

@@

#Add a title
plt.title("My AIA plot")
plt.show()
\end{minted}
\end{minipage}
\begin{minipage}{0.49\columnwidth}
\begin{minted}{python}
import sunpy.map
import matplotlib.pyplot as plt

#Read in a aia file using SunPy
aiamap = sunpy.map.Map('aia.fits')

@@

This design philosophy makes the behaviour of SunPy's visualisation 
routines intuitive to those who already understand how matplotlib's 
interface is designed as well as allowing the use of the standard 
matplotlib commands to change the plot, \textit{e.g.} set the title. 
More technically what is done in the plotting routines is firstly, to 
make sure that the current state of the figure and axes is mantined 
in the same way as it is inside the \texttt{pyplot} functions. For 
\texttt{Map} this means setting \texttt{pyplot.sci()} or set current 
image, so that \texttt{plt.colorbar()} 
automatically detects the plotted \texttt{AxesImage} object. 
Secondly, it also means returning the same type of object as the 
\texttt{pyplot} interfaces function, \textit{i.e.} for \texttt{Map} 
it returns the \texttt{AxesImage} instance which is the same type of 
return as \texttt{pyplot.imshow}.

>>>>

\section{Core Data Types}
The primary focus of the SunPy package is to provide data structures 
that are specifically designed for the various types of solar physics data, namely
images, time series, and spectra. SunPy provides three data core
objects to support these data types, 2D spatial data 
(\texttt{Map}), 1D temporal series (\texttt{LightCurve}) and 1 and 2D spectra 
(\texttt{Spectrum} and \texttt{Spectrogram}).

The philosophy of these objects is that they should provide a access to the original data
along with associated meta data and provide appropriate convenience functions to
enable data analysis and visualization. For all core data types, the data is stored in 
stored in the \texttt{.data}) attribute while the meta data is stored 
in the \texttt{.meta} attribute. It is 
possible to instantiate the data types from various different 
sources, \textit{e.g.} files, urls, arrays or time ranges.  In order to provide
instrument-specific specialisation, all the core SunPy data types support specialisation, 
\textit{i.e.} \texttt{Map} has an \texttt{AIAMap} sub-type. 
Data visualisation is provided by two functions; \texttt{peek()}, allows quick view plotting
, while more fine-grained control plotting that integrates with 
\texttt{matplotlib} is provided by \texttt{plot()}.

@@

\subsection{Visualisation}
\label{subsec:Viz}
As has been demonstrated in this section, the core SunPy datatypes 
include visualisation code that is specialised to that data type. 
These visualisation methods all currently utilise the matplotlib 
package, and are designed in such a way that they integrate well with 
the pyplot functional interface of matplotlib, as demonstrated in 
Figure \ref{code:mpl_1}.

%schriste - I vote for removing all of the following code as being too down in the weeds
\begin{listing}[H]
\begin{minipage}{0.49\columnwidth}
\begin{minted}[bgcolor=bg]{python}
import sunpy.map
import matplotlib.pyplot as plt

#Read in a aia file using SunPy
aiamap = SunPy.map.Map('aia.fits')

@@

#Add a title
plt.title("My AIA plot")
plt.show()
\end{minted}
\end{minipage}
\begin{minipage}{0.49\columnwidth}
\begin{minted}[bgcolor=bg]{python}
import SunPy.map
import matplotlib.pyplot as plt

#Read in a aia file using SunPy
aiamap = SunPy.map.Map('aia.fits')

@@

This design philosophy makes the behaviour of SunPy's visualisation 
routines intuitive to those who already understand how matplotlib's 
interface is designed as well as allowing the use of the standard 
matplotlib commands to change the plot, \textit{e.g.} set the title. 
More technically what is done in the plotting routines is firstly, to 
make sure that the current state of the figure and axes is mantained 
in the same way as it is inside the \texttt{pyplot} functions. For 
\texttt{Map} this means setting \texttt{pyplot.sci()} or set current 
image, so that \texttt{plt.colorbar()} 
automatically detects the plotted \texttt{AxesImage} object. 
Secondly, it also means returning the same type of object as the 
\texttt{pyplot} interfaces function, \textit{i.e.} for \texttt{Map} 
it returns the \texttt{AxesImage} instance which is the same type of 
return as \texttt{pyplot.imshow}.

### 2014-01-27 3-4-Helioviewer.tex
all of my changes

SunPy provides the ability to download images from Helioviewer Project 
servers.  The aim of the Helioviewer Project is to enable the exploration of 
solar and heliospheric data from multiple data sources (such as instrumentation 
and feature/event catalogs) via easy-to-use visual interfaces.  The primary 
data provided by the Helioviewer Project are images in the JPEG2000 format (for 
further details on the Helioviewer Project see \cite{muller2009} 
\footnote{SunPy is used in Helioviewer production servers to manage the 
download and ingestion of JPEG2000 files from remote servers.}.)  The JPEG2000 
files are typically highly compressed compared to the source FITS files they 
are generated from, and so can be a used as a shortcut to visualize large 
amounts of data from multiple sources.

The Helioviewer Project categorizes image data based on the physical
construction of the source instrument, using the following hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  The full list of image data sources available
can be found through the code snippet, Example
\ref{code:hv:datasources}.

\begin{listing}
\begin{minted}{python}
from sunpy.net.helioviewer import HelioviewerClient

hv = HelioviewerClient()
datasources = hv.get_data_sources()

# print a list of datasources and their associated ids
for observatory, instruments in datasources.items():
    for inst, detectors in instruments.items():
        for det, measurements in detectors.items():
            for meas, params in measurements.items():
                print("%s %s: %d" % (observatory, params['nickname'], 
                %%%%params['sourceId']))
\end{minted}
\caption{Initializing the Helioviewer client and listing the full
  range of image data available.}
\label{code:hv:datasources}
\end{listing}

@@

\begin{listing}
\begin{minted}{python}
hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]", x0=0, y0=0,
 width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

\begin{listing}
\begin{minted}{python}
hv.download_png('2099/01/01', 6, 
"[SDO,AIA,AIA,304,1,100],[SDO,AIA,AIA,193,1,50], \
[SOHO,LASCO,C2,white-light,1,100]", x0=0, y0=0, width=768, height=768)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_overlay_example}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

@@

\begin{listing}
\begin{minted}{python}
filepath = hv.download_jp2('2012/07/05 00:30:00', observatory='SDO',
 instrument='HMI', detector='HMI', measurement='continuum')
from sunpy.map import Map
hmi = Map(filepath)
hmi.submap([200,550],[-400,-200]).peek()
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy map object.}
\label{code:downloadjp2}
\end{listing}

>>>>

SunPy provides the ability to download images hosted by the Helioviewer Project.  
The aim of the Helioviewer Project is to enable the exploration of 
solar and heliospheric data from multiple data sources (such as instrumentation 
and feature/event catalogs) via easy-to-use visual interfaces.  The primary 
data provided by the Helioviewer Project are images in the JPEG2000 format (for 
further details on the Helioviewer Project see \cite{muller2009}). 
SunPy is used in Helioviewer production servers to manage the 
download and ingestion of JPEG2000 files from remote servers.  The JPEG2000 
files are typically highly compressed compared to the source FITS files they 
are generated from, and so can be a used as a shortcut to visualize large 
amounts of data from multiple sources.

The Helioviewer Project categorizes image data based on the physical
construction of the source instrument, using the following hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  The full list of image data sources available
can be found through the code snippet, Example
\ref{code:hv:datasources}.
%schriste - what is the point of the following listing? remove?
\begin{listing}[H]
\begin{minted}{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> datasources = hv.get_data_sources()

# print a list of datasources and their associated ids
>>> for observatory, instruments in datasources.items():
>>>     for inst, detectors in instruments.items():
>>>         for det, measurements in detectors.items():
>>>             for meas, params in measurements.items():
>>>                 print(observatory, params['nickname'], 
>>>                 params['sourceId']))
\end{minted}
\caption{Initializing the Helioviewer client and listing the full
  range of image data available.}
\label{code:hv:datasources}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}{python}
hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]", x0=0, y0=0,
 width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}{python}
hv.download_png('2099/01/01', 6, 
"[SDO,AIA,AIA,304,1,100],[SDO,AIA,AIA,193,1,50], \
[SOHO,LASCO,C2,white-light,1,100]", x0=0, y0=0, width=768, height=768)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_overlay_example}
\end{center}
\caption{Acquisition of a PNG image composed from data from three
  separate sources.}
\label{code:hv:overlaid}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}{pycon}
>>> filepath = hv.download_jp2('2012/07/05 00:30:00', observatory='SDO',
>>>                            instrument='HMI', detector='HMI', 
>>>                            measurement='continuum')
>>> from sunpy.map import Map
>>> hmi = Map(filepath)
>>> hmi.submap([200,550],[-400,-200]).peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{helioviewer_hmi_continuum_jp2_to_map}
\end{center}
\caption{Acquisition and display of a Helioviewer JPEG2000 file as a
  SunPy map object.}
\label{code:downloadjp2}
\end{listing}
